Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de~Rham complex over convex domains and domains with shellable triangulations

### 2025-02-12 main.tex
tweaks and language

\newcommand{\mwl}[1]{{\color{red}#1}}
\newcommand{\notice}[1]{{\color{red}REMARK: #1}}
\newcommand{\disk}{\cye{ball}}
\newcommand{\todo}[1]{{\color{RedOrange}\textbf{#1}}}

@@

\begin{abstract}
    We construct potentials for the gradient, the curl, and the divergence operators over domains with shellable triangulations. 
    Notably, the class of shellable triangulations includes local patches (stars) in two or three dimensions. 
    The operator norms of our potentials satisfy explicitly computable bounds that depend only on the geometry. 
    We thus obtain computable upper bounds for constants in Poincar\'e--Friedrichs inequalities and computable lower bounds for the eigenvalues of vector Laplacians. 
    % We derive computable bound on the operator norms of our potentials that only depend on the geometry. 
    % 
    %     We construct upper bounds for Poincar\'e--Friedrichs constants over triangulated domains. 
    %     This includes lower estimates for the singular values of the gradient, curl, and divergence operators as special cases. 
    %     The main application that we envision are computable estimates over finite element patches. 
    %
    As an additional result with independent standing, we establish Poincar\'e--Friedrichs inequalities with computable constants for the $L^{p}$ de~Rham complex over bounded convex domains. 
    This is achieved via regularized Poincar\'e and Bogovski\u{\i} potential operators whose operator norms we bound.
    We express all our main results in the calculus of differential forms and treat the gradient, curl, and divergence operators as its particular instances. 
    Computational examples illustrate the theoretical findings.
\end{abstract}

@@

Potentials for the differential operators of vector calculus and exterior calculus are of fundamental importance. 
The operator norms of these potentials are upper bounds for the Poincar\'e--Friedrichs constants.
These quantify the fundamental stability properties of {numerous} partial differential equations % DONE
and enter the stability and convergence theory of numerical methods. 
Upper bounds of the Poincar\'e--Friedrichs constants also provide lower bounds of the eigenvalues of the associated Laplacians.
% DONE {and the eigenvalues of finite element matrices. (doubtful statement?)}
However, while potentials and Poincar\'e--Friedrichs constants for the gradient have been subject to extensive study,
quantifiable results regarding the curl and divergence operators, or more generally the exterior derivative, are largely unavailable.

This manuscript contributes to the theory of computable estimates for Poincar\'e--Friedrichs inequalities for the differential operators of vector calculus. 
How to construct potentials for the gradient over triangulated domains is well-documented in the literature. 
Here, we extend this construction to the curl and divergence operators, as well as the exterior derivative. 
However, to proceed in the general exterior derivative case, we restrict our efforts to so-called shellable triangulations. 
Only contractible domains can ever admit a shellable triangulation, but having computable upper bounds for such domains is an important stepping stone towards more general situations. 
The class of shellable triangulations includes practically relevant triangulations: 
for example, local triangulations around a simplex within a larger triangulation (the so-called local patches or stars) are shellable in dimensions two and three; see also Figure~\ref{figure:patches}. 
% DONE {[TCF: Is it really true if the simplex is of highest dimension, e.g., a tetrahedron in 3D?]} 
Additionally, we include a study of regularized Poincar\'e and Bogovski\u{\i} operators that leads to new Poincar\'e--Friedrichs inequalities with computable constants for the whole $L^{p}$ de~Rham complex over bounded convex domains.

@@

We are interested in a constant $C_{\grad,\Omega,p} > 0$ such that the following holds:
for every gradient vector field $\bff \in \nabla W^{1,p}(\Omega)$ 
there exists a scalar potential $u \in W^{1,p}(\Omega)$
such that $\nabla u = \bff$ and 
\begin{align}\label{math:intro:inequalitypf:gradient} 
    C_{\grad,\Omega,p} 
    \| \bff \|_{L^{p}(\Omega)}
This inequality is called \emph{Poincar\'e--Friedrichs inequality} and the constant $C_{\grad,\Omega,p}$ is called the \emph{Poincar\'e--Friedrichs constant} with exponent $p$. 
The question is therefore whether we can always find a gradient potential of sufficiently small norm so this inequality holds. 
One possible choice is the norm-minimizing potential 
\begin{align}\label{math:intro:gradientpotential}
    \Phi_{\grad}( \bff ) := \argmin\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \nabla u = \bff } } \| u \|_{L^{p}(\Omega)}.
In the present setting, where $\Omega$ is connected and the potential of the different gradient potentials can thus only differ by a constant, computing $\Phi_{\grad}( \bff )$ is a one-dimensional convex minimization problem. 
This inequality is therefore equivalent to  
\begin{align}\label{math:intro:poincarefriedrichs:grad}
    \min\limits_{c\in\bbR}
    \| u - c \|_{L^{p}(\Omega)}
    C_{\grad,\Omega,p} \| \nabla u \|_{L^{p}(\Omega)},
    u \in W^{1,p}(\Omega) 
    .\tag{PF} 
Such an inequality holds if we can bound the \emph{operator norm} of this gradient potential operator, and then the best constant is  
%the $L^{p}(\Omega)$ norm of the potential $\Phi_{\grad}( \bff )$ is bounded in terms of the $\bfL^{p}(\Omega)$ norm of the source field $\bff$.
%Hence, we want to evaluate the \emph{operator norm} of the potential operator 
\begin{align}\label{math:intro:gradient:potentialnorm}
    C_{\grad,\Omega,p} := \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \Phi_{\grad}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\fooL^{p}(\Omega)} }.
Finding the norm-minimizing potential and the optimal Poincar\'e--Friedrichs constant is generally difficult. 
Attention has been given to \emph{linear} potentials instead of the norm-minimizing nonlinear potential~\eqref{math:intro:gradientpotential}. 
The operator norm of any linear potential construction serves as an upper bound for the Poincar\'e--Friedrichs constant. 
One very simple example for such a bounded linear potential operator is the average-free potential, 
\begin{align}\label{math:intro:poincarepotential}
    \Phi_{\varnothing}( \bff ) 
    \argmin\limits_{ \substack{ v \in W^{1,p}(\Omega) \\ \int_\Omega v = 0 } } \| \nabla v - \bff \|_{L^{p}(\Omega)},
    \bff \in \nabla W^{1,p}(\Omega)
where the mean value of $u$ is fixed to zero. 
Its operator norm is the Poincar\'e constant $C_{\varnothing,\Omega,p} > 0$ that satisfies % for the \emph{Poincar\'e inequality} 
\begin{align}\label{math:intro:poincare:grad}
    C_{\varnothing,\Omega,p} \| \nabla u \|_{L^{p}(\Omega)},
    u \in W^{1,p}(\Omega)
    , \tag{P} 
where $u_\Omega$ denotes the average of $u$. We emphasize that the average-free potential is generally not the norm-minimizing potential unless $p = 2$.
Hence, the optimal Poincar\'e constant and the optimal Poincar\'e--Friedrichs differ in general. 
Considerable research efforts have gone into determining constants in the Poincar\'e--Friedrichs inequalities~\eqref{math:intro:poincarefriedrichs:grad} or Poincar\'e inequalities~\eqref{math:intro:poincare:grad}. Upper estimates for these constants also correspond to lower bounds for the spectra of Neumann--Laplacians over those domains. 
The relationship between the Poincar\'e constant and the Poincar\'e--Friedrichs constant will be elaborated upon in later sections of this manuscript.

@@

We study potentials and Poincar\'e--Friedrichs inequalities for the curl or divergence operators in vector calculus.
There are substantial changes and much fewer results are available in the literature. 
We use the spaces of vector field 
\begin{align}
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in L^{p}(\Omega)^{3} : \curl \bfu \in L^{p}(\Omega)^{3} \right\}
    \label{math:intro:Wcurl}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in L^{p}(\Omega)^{3} : \divergence \bfu \in L^{p}(\Omega) \right\}
    \label{math:intro:Wdiv}
    .
\end{align}
Their members are those vector fields in Lebesgue spaces whose distributional curls and divergences, respectively, are in Lebesgue spaces. 
In contrast to the gradient, this only requires that certain sums of distributional partial derivatives are integrable, 
and hence these spaces are not classical Sobolev spaces of vector fields. 
Our objective is to find bounded potentials (i.e., right inverses) for the operators 
\begin{align*}
    \curl : \bfW^{p}(\curl,\Omega) \rightarrow L^{p}(\Omega)^{3},
    \qquad 
    \divergence : \bfW^{p}(\divergence,\Omega) \rightarrow L^{p}(\Omega),
\end{align*}
We are interested in the natural analogues for the Poincar\'e--Friedrichs inequality of the gradient~\eqref{math:intro:poincarefriedrichs:grad},
which for the curl and divergence are the inequalities 
\begin{align}
    \min\limits_{ \substack{ \bfv \in \bfW^{p}(\curl,\Omega) \\ \curl \bfv = \bff } } 
    \| \bfv \|_{L^{p}(\Omega)}
    &\leq 
    C_{\curl,\Omega,p}
    \| \bff \|_{L^{p}(\Omega)}
    \label{math:intro:Wcurl:pf}
    ,
    \\%\qquad 
    \min\limits_{ \substack{ \bfv \in \bfW^{p}(\divergence,\Omega) \\ \divergence \bfv = f } } 
    \| \bfv \|_{L^{p}(\Omega)}
    &\leq 
    C_{\divergence,\Omega,p}
    \| f \|_{L^{p}(\Omega)}
    \label{math:intro:Wdiv:pf}
    .
\end{align}
The fundamental difference to the gradient is that the curl and divergence have infinite-dimensional kernels. 
The kernel of the gradient is the one-dimensional space of constant functions, and it is thus trivially complemented for all $p$, with a canonical choice of projection. 
By contrast, the kernels of the curl and divergence operators are generally infinite-dimensional. 
Moreover, it is not immediately evident that the kernels of the curl and divergence operators are complemented in the Banach space case when $p \neq 2$, 
and a canonical projection only exists here in the Hilbert setting $p=2$. 
In that sense, there generally is no natural analogue to the Poincar\'e inequality~\eqref{math:intro:poincare:grad} for the curl and divergence.

@@

Though we present our results in vector calculus, our main arguments are given in the formalism of exterior calculus. 
Exterior calculus~\cite{greub1967multilinear,lee2012smooth} is used ubiquitously in the mathematical literature of physics and engineering and has found widespread adoption in the theoretical and numerical analysis for vector-valued partial differential equations~\cite{hiptmair2002finite, gross2004electromagnetic, arnold2006finite, arnold2009geometric, arnold2010finite, demlow2014posteriori, licht2021local, arnold2021complexes}. 
This formalism is independent of the spatial dimension and highlights the underlying geometric structures common to the gradient, curl, and divergence operators in three space dimensions.
For every $u \in W^{p}\Alt^{k}(\Omega)$, there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan w = \cartan u$ and such that (all the notation is fixed in detail in Section~\ref{section:calculus} later)
\begin{align} \label{math:intro:exterior:pf} 
    \| u \|_{L^{p}(\Omega)} \leq  C_{k,\Omega,p} \| \cartan w \|_{L^{p}(\Omega)}.
\end{align} 
For the purpose of our discussion, this formalism allows us to leverage results from a larger body of literature in differential geometry and functional analysis. 
% This formalism is independent of the spatial dimension and highlights underlying geometric structures common to the gradient, curl, and divergence operators.
% In this setting, the common writing of the Poincar\'e--Friedrichs inequalities we are interested in is just as~\eqref{math:PF_gradient_1}:

@@

The qualitative existence of Poincar\'e--Friedrichs inequalities for the differential operators of vector calculus and exterior calculus is known for a large class of domains. 
There is a particularly large body of literature on the Poincar\'e inequality for the gradient and its numerous variants,
which may include weighted integrals or boundary terms. 
These Poincar\'e-type inequalities are often derived via non-constructive arguments, such as the Rellich embedding theorem, 
and thus the resulting constants are not explicitly computable. 
% DONT {[TCF: I believe this is often referred to as the Tartare-Petree theorem]}

@@

% curl 
The core challenges can be found in the discussion of the curl operator in three dimensions,
where Poincar\'e--Friedrichs inequalities have appeared under different names. 
Let us assume for a moment that $\Omega$ is a weakly Lipschitz domain with trivial topology, and consider only the Hilbert case $p=2$. 
Then the constant in \eqref{math:intro:Wcurl:pf} agrees with the constant in the so-called Poincar\'e--Friedrichs--Weber inequality
\begin{align}
    \label{math:intro:weber}
    \| \bfu\|_{L^{2}(\Omega)}   \leq C_{\curl,\Omega,p} \| \curl \bfu \|_{L^{2}(\Omega)},
\end{align}
valid for all $\bfu \in \bfW^{2}(\curl,\Omega) \cap \bfW^{2}(\divergence,\Omega)$ that satisfy $\divergence \bfu = 0$ and have vanishing normal or vanishing tangential trace along $\partial \Omega$. 
Equivalently, \eqref{math:intro:weber} is valid for all $\bfu \in \bfW^{2}(\curl,\Omega)$ that are $L^{2}$-orthogonal to the gradients of scalar fields in $W^{1,2}(\Omega)$
or that have vanishing tangential trace and are $L^{2}$-orthogonal to the gradients of those scalar fields in $W^{1,2}(\Omega)$ that satisfy Dirichlet boundary conditions.
% Yet another rewriting is 
% \label{math:PF_orth} \begin{align}
%     \| \bfu\|_{L^{2}(\Omega)} & \leq C_{\curl,\Omega,2} \| \curl \bfu \|_{L^{2}(\Omega)}, 
%     \label{math:PF_orth_curl}
%     \\
%     \| \bfu\|_{L^{2}(\Omega)} & \leq \tilde C_{\divergence,\Omega,2} \| \divergence \bfu \|_{L^{2}(\Omega)}, 
%     \label{math:PF_orth_div}
% \end{align}
% where the former holds for all vector fields $\bfu \in \bfW^{2}(\curl,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv \in \bfW^{2}(\curl,\Omega)$ with $\curl \bfv = 0$,
% and where the latter holds for all functions $\bfu \in \bfW^{2}(\divergence,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv \in \bfW^{2}(\divergence,\Omega)$ with $\divergence \bfv = 0$. 
We refer to Equation~(5) in~\cite{Fried_diff_forms_55}, Equation~(2) in \cite{Gaff_Hilbert_harm_55}, \cite{Web_compact_Maxw_80} as well as \cite[Lemmas~3.4 and~3.6]{Gir_Rav_NS_86}, \cite[Proposition~7.4]{Fer_Gil_Maxw_BC_97}, and the references therein. 
The general case of $L^p$ differential forms over Lipschitz manifolds subject to partial boundary conditions is discussed in~\cite{goldshtein2011hodge}.
However, while many of the above results rely on non-constructive estimates, 
we are interested in practically computable upper bounds.

@@

When the domain is not convex but star-shaped with respect to a ball,
then several estimates for Poincar\'e--Friedrichs constants are known.\footnote{In fact, Poincar\'e--Friedrichs inequalities even hold over any star-shaped open bounded set; see~\cite[Theorem~3.1]{hurri1988poincare}.}
% DONE {[TCF: I don't exactly get the point of this sentence.].}
Polynomial interpolation estimates already imply the gradient Poincar\'e--Friedrichs inequality~\cite{brenner2008mathematical,ern2021finite}. 
We pay particular attention to the regularized Poincar\'e and Bogovski\u{\i} potential operators for the exterior derivative,
such as those of Costabel and McIntosh~\cite{costabel2010bogovskiui}.
In principle, the operator norms of those potentials as mapping between Lebesgue spaces of differential forms 
are upper estimates for the Poincar\'e--Friedrichs constants of the domain. 
Here, estimates for the higher-order seminorms of these potentials are available in~\cite{guzman2021estimation}, 
but estimates in Lebesgue norms have not been made explicit in the literature yet, to the best of our knowledge.
We particularly emphasize that these estimates for domains star-shaped with respect to a ball have in common that they rely on upper bounds for the \emph{eccentricity} of the domain.

@@

Let us briefly discuss the practical limitations of the aforementioned estimates for Poincar\'e--Friedrichs constants for convex domains or domains that are star-shaped with respect to a ball. 
Recall that the main objective of this manuscript is bounding Poincar\'e--Friedrichs constants over domains with shellable triangulations, with a key application being local stars within triangulated domains. 
While not all local stars describe convex subdomains, most local stars are star-shaped with respect to a ball. 
Even though that would enable, e.g., the averaged Poincar\'e and Bogovski\u{\i} operators~\cite{costabel2010bogovskiui}, 
the estimates that rely on this geometric condition deteriorate when the aforementioned ball has radius much smaller than the domain diameter. 
This would not be as much a problem over local patches (stars) around interior subsimplices, where the size of the interior ball only depends on the shape regularity of the triangulation. 
But the interior ball can be arbitrarily small when the local patch is around a boundary simplex, even if the mesh has good shape regularity: 
this occurs most prominently when the domain has sharp reentrant corners. Some illustrative limit cases include the slit domain~\cite{veeser2012poincare} and the crossed bricks domain~\cite{licht2019smoothed}, 
which contain local finite element patches that are not star-shaped with respect to any ball. 
In view of this, we refrain from treating local patches (stars) in triangulations as domains star-shaped with respect to a ball.

@@

Geometric settings that admit finite triangulations enable different pathways to obtain Poincar\'e--Friedrichs inequalities. 
We review some of the main outcomes.

>>>>

@@

\begin{abstract}
    We construct potentials for the gradient, the curl, and the divergence operators over domains with shellable triangulations.
    Notably, the class of shellable triangulations includes local patches (stars) in two or three dimensions.
    The operator norms of our potentials satisfy explicitly computable bounds that depend only on the geometry.
    We thus compute upper bounds for constants in Poincar\'e--Friedrichs inequalities and lower bounds for the eigenvalues of vector Laplacians.
    % We derive computable bound on the operator norms of our potentials that only depend on the geometry.
    % 
    %     We construct upper bounds for Poincar\'e--Friedrichs constants over triangulated domains. 
    %     This includes lower estimates for the singular values of the gradient, curl, and divergence operators as special cases. 
    %     The main application that we envision are computable estimates over finite element patches. 
    %
    As an additional result with independent standing, we establish Poincar\'e--Friedrichs inequalities with computable constants for the $L^{p}$ de~Rham complex over bounded convex domains, derived as explicit operator norms of regularized Poincar\'e and Bogovski\u{\i} potential operators.
    We express all our main results in the calculus of differential forms and treat the gradient, curl, and divergence operators as instances of the exterior derivative.
    Computational examples illustrate the theoretical findings.
\end{abstract}

@@

Potentials for the differential operators of vector calculus and exterior calculus are of fundamental importance.
The operator norms of these potentials are upper bounds for the Poincar\'e--Friedrichs constants.
These quantify the fundamental stability properties of {numerous} partial differential equations % DONE
and enter the stability and convergence theory of numerical methods.
Upper bounds of the Poincar\'e--Friedrichs constants also provide lower bounds of the eigenvalues of the associated Laplacians.
% DONE {and the eigenvalues of finite element matrices. (doubtful statement?)}
However, while potentials and Poincar\'e--Friedrichs constants for the gradient have been subject to extensive study,
quantifiable results regarding the curl and divergence operators, or more generally, the exterior derivative, are largely unavailable.

This manuscript contributes to the theory of computable estimates for Poincar\'e--Friedrichs inequalities for the differential operators of vector calculus.
How to construct potentials for the gradient over triangulated domains is well-documented in the literature.
Here, we extend this construction to the curl and divergence operators, treating them as instances of the exterior derivative.
However, to proceed in the general exterior derivative case, we restrict our efforts to so-called shellable triangulations.
The class of shellable triangulations includes practically relevant triangulations:
for example, local triangulations around simplices within a larger triangulation (the so-called local patches or stars) are shellable in dimensions two and three; see also Figure~\ref{figure:patches}.
Only contractible domains can ever admit a shellable triangulation, but having computable upper bounds for such domains is an important stepping stone toward more general situations.
Additionally, we include a study of regularized Poincar\'e and Bogovski\u{\i} operators that leads to new Poincar\'e--Friedrichs inequalities with computable constants for the whole $L^{p}$ de~Rham complex over bounded convex domains.
% DONE {[TCF: Is it really true if the simplex is of highest dimension, e.g., a tetrahedron in 3D?]}

@@

We are interested in a constant $C_{\grad,\Omega,p} > 0$ such that the following holds:
for every gradient vector field $\bff \in \nabla W^{1,p}(\Omega)$ 
there exists a scalar potential $u \in W^{1,p}(\Omega)$
such that $\nabla u = \bff$ and 
\begin{align}\label{math:intro:inequalitypf:gradient} 
    \| u \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} 
    \| \bff \|_{L^{p}(\Omega)}
    .
\end{align}
This inequality is called \emph{Poincar\'e--Friedrichs inequality} and the constant $C_{\grad,\Omega,p}$ is called the \emph{Poincar\'e--Friedrichs constant} with exponent $p$. 
The question is therefore whether we can always find a gradient potential of sufficiently small norm so this inequality holds. 
One possible choice is the norm-minimizing potential 
\begin{align}\label{math:intro:gradientpotential}
    \Phi_{\grad}( \bff ) := \argmin\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \nabla u = \bff } } \| u \|_{L^{p}(\Omega)}.
\end{align}
In the present setting, where $\Omega$ is connected and the potential of the different gradient potentials can thus only differ by a constant, computing $\Phi_{\grad}( \bff )$ is a one-dimensional convex minimization problem. 
This inequality is therefore equivalent to  
\begin{align}\label{math:intro:poincarefriedrichs:grad}
    \min\limits_{c\in\bbR}
    \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} \| \nabla u \|_{L^{p}(\Omega)},
    \quad 
    \forall
    u \in W^{1,p}(\Omega) 
    .\tag{PF} 
\end{align}
Such an inequality holds if we can bound the \emph{operator norm} of this gradient potential operator, and then the best constant is  
%the $L^{p}(\Omega)$ norm of the potential $\Phi_{\grad}( \bff )$ is bounded in terms of the $\bfL^{p}(\Omega)$ norm of the source field $\bff$.
%Hence, we want to evaluate the \emph{operator norm} of the potential operator 
\begin{align}\label{math:intro:gradient:potentialnorm}
    C_{\grad,\Omega,p} := \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \Phi_{\grad}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\fooL^{p}(\Omega)} }.
\end{align}
Finding the norm-minimizing potential and the optimal Poincar\'e--Friedrichs constant is generally difficult. 
Attention has focused on \emph{linear} potentials instead of the norm-minimizing nonlinear potential~\eqref{math:intro:gradientpotential}.
The operator norm of any linear potential construction serves as an upper bound for the Poincar\'e--Friedrichs constant. 
One straightforward example for such a bounded linear potential operator is the average-free potential,
\begin{align}\label{math:intro:poincarepotential}
    \Phi_{\varnothing}( \bff ) 
    := 
    \argmin\limits_{ \substack{ v \in W^{1,p}(\Omega) \\ \int_\Omega v = 0 } } \| \nabla v - \bff \|_{L^{p}(\Omega)},
    \quad 
    \forall 
    \bff \in \nabla W^{1,p}(\Omega)
    ,
\end{align} 
whose mean value is fixed to zero.
Its operator norm is the Poincar\'e constant $C_{\varnothing,\Omega,p} > 0$ that satisfies % for the \emph{Poincar\'e inequality} 
\begin{align}\label{math:intro:poincare:grad}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{\varnothing,\Omega,p} \| \nabla u \|_{L^{p}(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    , \tag{P} 
\end{align}
where $u_\Omega$ denotes the average of $u$. We emphasize that the average-free potential is generally not the norm-minimizing potential unless $p = 2$.
Hence, the optimal Poincar\'e constant and the optimal Poincar\'e--Friedrichs generally differ.
Considerable research efforts have gone into determining constants in the Poincar\'e--Friedrichs inequalities~\eqref{math:intro:poincarefriedrichs:grad} or Poincar\'e inequalities~\eqref{math:intro:poincare:grad}. Upper estimates for these constants also correspond to lower bounds for the spectra of Neumann--Laplacians over those domains.
The relationship between the Poincar\'e constant and the Poincar\'e--Friedrichs constant will be elaborated upon in later sections of this manuscript.

@@

We study potentials and Poincar\'e--Friedrichs inequalities for the curl or divergence operators in vector calculus.
There are substantial changes and much fewer results are available in the literature. 
We use the spaces of vector field 
\begin{align}
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in L^{p}(\Omega)^{3} : \curl \bfu \in L^{p}(\Omega)^{3} \right\}
    \label{math:intro:Wcurl}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in L^{p}(\Omega)^{3} : \divergence \bfu \in L^{p}(\Omega) \right\}
    \label{math:intro:Wdiv}
    .
\end{align}
Their members are those vector fields in Lebesgue spaces whose distributional curls and divergences, respectively, are in Lebesgue spaces. 
In contrast to the gradient, this only requires that certain sums of distributional partial derivatives are integrable, 
and hence these spaces are not classical Sobolev spaces of vector fields. 
Our objective is to find bounded potentials (i.e., right inverses) for the operators 
\begin{align*}
    \curl : \bfW^{p}(\curl,\Omega) \rightarrow L^{p}(\Omega)^{3},
    \qquad 
    \divergence : \bfW^{p}(\divergence,\Omega) \rightarrow L^{p}(\Omega).
\end{align*}
We are interested in the natural analogues for the Poincar\'e--Friedrichs inequality of the gradient~\eqref{math:intro:poincarefriedrichs:grad},
which for the curl and divergence are the inequalities 
\begin{align}
    \min\limits_{ \substack{ \bfv \in \bfW^{p}(\curl,\Omega) \\ \curl \bfv = \bff } } 
    \| \bfv \|_{L^{p}(\Omega)}
    &\leq 
    C_{\curl,\Omega,p}
    \| \bff \|_{L^{p}(\Omega)}
    \label{math:intro:Wcurl:pf}
    ,
    \\%\qquad 
    \min\limits_{ \substack{ \bfv \in \bfW^{p}(\divergence,\Omega) \\ \divergence \bfv = f } } 
    \| \bfv \|_{L^{p}(\Omega)}
    &\leq 
    C_{\divergence,\Omega,p}
    \| f \|_{L^{p}(\Omega)}
    \label{math:intro:Wdiv:pf}
    .
\end{align}
The fundamental difference to the gradient is that the curl and divergence have infinite-dimensional kernels.
The kernel of the gradient is the one-dimensional space of constant functions, and it is thus trivially complemented for all $p$, with a canonical choice of projection.
By contrast, the kernels of the curl and divergence operators are generally infinite-dimensional.
Moreover, it is not immediately evident that the kernels of the curl and divergence operators are complemented in the Banach space case when $p \neq 2$,
and a canonical projection only exists here in the Hilbert setting $p=2$.
In that sense, there generally is no natural analog to the Poincar\'e inequality~\eqref{math:intro:poincare:grad} for the curl and divergence.

@@

Though we present our results in vector calculus, our main arguments are given in the formalism of exterior calculus. 
Exterior calculus~\cite{greub1967multilinear,lee2012smooth} is used ubiquitously in the mathematical literature of physics and engineering and has found widespread adoption in the theoretical and numerical analysis for vector-valued partial differential equations~\cite{hiptmair2002finite, gross2004electromagnetic, arnold2006finite, arnold2009geometric, arnold2010finite, demlow2014posteriori, licht2021local, arnold2021complexes}. 
This formalism is independent of the spatial dimension and highlights the underlying geometric structures common to the gradient, curl, and divergence operators in three space dimensions.
For every $u \in W^{p}\Alt^{k}(\Omega)$, there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan w = \cartan u$ and such that (all the notation is fixed in detail in Section~\ref{section:calculus} later)
\begin{align}\label{math:intro:exterior:pf} 
    \| w \|_{L^{p}(\Omega)} \leq  C_{k,\Omega,p} \| \cartan w \|_{L^{p}(\Omega)}.
\end{align} 
For the purpose of our discussion, this formalism allows us to leverage results from a larger body of literature in differential geometry and functional analysis. 
% This formalism is independent of the spatial dimension and highlights underlying geometric structures common to the gradient, curl, and divergence operators.
% In this setting, the common writing of the Poincar\'e--Friedrichs inequalities we are interested in is just as~\eqref{math:PF_gradient_1}:

@@

The qualitative existence of Poincar\'e--Friedrichs inequalities for the differential operators of vector calculus and exterior calculus is known for a large class of domains.
There is a vast body of literature on the Poincar\'e inequality for the gradient and its numerous variants,
which may include weighted integrals or boundary terms.
We emphasize that non-constructive arguments, such as the Rellich embedding theorem, generally do not yield explicitly computable constants.
% DONT {[TCF: I believe this is often referred to as the Tartare-Petree theorem]}

@@

% curl 
The core challenges can be found in the discussion of the curl operator in three dimensions,
where Poincar\'e--Friedrichs inequalities have appeared under different names. 
Let us assume for a moment that $\Omega$ is a weakly Lipschitz domain with trivial topology, and consider only the Hilbert case $p=2$. 
Then the constant in~\eqref{math:intro:Wcurl:pf} agrees with the constant in the so-called Poincar\'e--Friedrichs--Weber inequality
\begin{align}
    \label{math:intro:weber}
    \| \bfu\|_{L^{2}(\Omega)}   \leq C_{\curl,\Omega,p} \| \curl \bfu \|_{L^{2}(\Omega)},
\end{align}
valid for all $\bfu \in \bfW^{2}(\curl,\Omega) \cap \bfW^{2}(\divergence,\Omega)$ that satisfy $\divergence \bfu = 0$ and have vanishing normal or vanishing tangential trace along $\partial \Omega$. 
Equivalently,~\eqref{math:intro:weber} is valid for all $\bfu \in \bfW^{2}(\curl,\Omega)$ that are $L^{2}$-orthogonal to the gradients of scalar fields in $W^{1,2}(\Omega)$
or that have vanishing tangential trace and are $L^{2}$-orthogonal to the gradients of those scalar fields in $W^{1,2}(\Omega)$ that satisfy Dirichlet boundary conditions.
% Yet another rewriting is 
% \label{math:PF_orth} \begin{align}
%     \| \bfu\|_{L^{2}(\Omega)} & \leq C_{\curl,\Omega,2} \| \curl \bfu \|_{L^{2}(\Omega)}, 
%     \label{math:PF_orth_curl}
%     \\
%     \| \bfu\|_{L^{2}(\Omega)} & \leq \tilde C_{\divergence,\Omega,2} \| \divergence \bfu \|_{L^{2}(\Omega)}, 
%     \label{math:PF_orth_div}
% \end{align}
% where the former holds for all vector fields $\bfu \in \bfW^{2}(\curl,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv \in \bfW^{2}(\curl,\Omega)$ with $\curl \bfv = 0$,
% and where the latter holds for all functions $\bfu \in \bfW^{2}(\divergence,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv \in \bfW^{2}(\divergence,\Omega)$ with $\divergence \bfv = 0$. 
We refer to Equation~(5) in~\cite{Fried_diff_forms_55}, Equation~(2) in~\cite{Gaff_Hilbert_harm_55},~\cite{Web_compact_Maxw_80} as well as~\cite[Lemmas~3.4 and~3.6]{Gir_Rav_NS_86},~\cite[Proposition~7.4]{Fer_Gil_Maxw_BC_97}, and the references therein. 
The general case of $L^p$ differential forms over Lipschitz manifolds subject to partial boundary conditions is discussed in~\cite{goldshtein2011hodge}.
However, while many of the above results rely on non-constructive estimates, 
we are interested in practically computable upper bounds.

@@

When the domain is not convex but star-shaped with respect to a ball,
then several estimates for Poincar\'e--Friedrichs constants are known.\footnote{In fact, Poincar\'e--Friedrichs inequalities even hold over any star-shaped open bounded set; see~\cite[Theorem~3.1]{hurri1988poincare}.}
% DONE {[TCF: I don't exactly get the point of this sentence.].}
Polynomial interpolation estimates already imply the gradient Poincar\'e--Friedrichs inequality~\cite{brenner2008mathematical,ern2021finite}.
We pay particular attention to the regularized Poincar\'e and Bogovski\u{\i} potential operators for the exterior derivative,
such as those of Costabel and McIntosh~\cite{costabel2010bogovskiui}.
In principle, if we conceive these potentials as mappings between Lebesgue spaces of differential forms, then their operator norms are upper estimates for the Poincar\'e--Friedrichs constants of the domain.
We are aware of estimates for the higher-order seminorms of these potentials~\cite{guzman2021estimation},
but estimates in Lebesgue norms have not been made explicit in the literature yet, to the best of our knowledge.
We particularly emphasize that all these estimates for domains star-shaped with respect to a ball have in common that they rely on upper bounds for the \emph{eccentricity} of the domain.

@@

Let us briefly discuss the practical limitations of the aforementioned estimates for Poincar\'e--Friedrichs constants for convex domains or domains that are star-shaped with respect to a ball.
Recall that the main objective of this manuscript is bounding Poincar\'e--Friedrichs constants over domains with shellable triangulations, with a key application being local stars within triangulated domains.
Not all local stars describe convex subdomains.
Even if the local stars are star-shaped with respect to a ball,
which would enable, e.g., the averaged Poincar\'e and Bogovski\u{\i} operators~\cite{costabel2010bogovskiui},
the estimates that rely on this geometric condition deteriorate when the aforementioned ball has radius much smaller than the domain diameter.
While this is not as much a problem over local patches (stars) around interior subsimplices, where the size of the interior ball only depends on the shape regularity of the triangulation, the interior ball can be arbitrarily small when the local patch is around a boundary simplex, even if the triangulation displays decent shape regularity.
This occurs most prominently when the domain has sharp reentrant corners, illustrative limit cases including the slit domain~\cite{veeser2012poincare} and the crossed bricks domain~\cite{licht2019smoothed},
which contain local finite element patches that are not star-shaped with respect to any ball.
In view of this, we refrain from treating local patches (stars) in triangulations as domains star-shaped with respect to a ball.

@@

Geometric settings that admit finite triangulations enable different pathways to obtain Poincar\'e--Friedrichs inequalities.
We review the main outcomes.
